Exploring the Usefulness of Varactor loaded Nonlinear Transmission Lines

### 2020-06-10 RohdeAndSchwarzWhitePaper/Draft_White_Paper.tex
Fixed negative Resistance in doc and added a sim file for risetime
Ran report through grammarly and rebuilt
Next step is to incorporate edits made by Mr. Beck and Mr. Fetche

First, we discuss the necessity of non-linear components in such a Nonlinear
Transmission line or \emph{NLTL} for short. We then discuss ways in which the
devices we use in simulation are modeled and the obstacles we faced in
constructing these models. We simulate these nonlinear lines using MICROCAP and
discuss possible reasons for the discrepancies we have seen between research and
simulations.

@@

Much like a wave travelling in the ocean towards the shore, certain sections of
the wave travel at different speeds despite the travel of the wave as a whole.
This can be seen first hand by a surfer utilizing their position on the wave to
have the right speed and stability.

In electronics this same concept is present in optical fibers or waveguides. As
we reflect a signal beam into a guide, the phase velocity can actually exceed
the speed of light, however the speed of the the information characterized by
the group velocity is what we care about.

\subsection{Types of NLTL waves}\label{sub:NLTLTypes}
While treating the physics of non-linear waves is beyond the scope of this
paper, the use of NLTL's in transmission lines is founded upon the existence of
waves that do not behave like the pure sine waves we usually consider in
electronics. These both types of waves can be present and used in techonology to
lessen the time it takes for an input waveform to reach its final value as well
as lessen the time to return to its ``off'' voltage state. In other words, these
can be used to sharpen both rising and falling edges of a square
wave input\cite{afshari2005nonlinear}.

Informally, \emph{Solitons} are much like the type of wave we see
travelling in the open ocean without breaking and are characterized by their
ability to interact with other waves without losing energy. This means that
after passing another wave it largely maintains its shape and is only shifted in
phase (experiences time delay).

@@

Another main concept of operation of these types of circuits, is the need for
non-linearity to outweigh possible wave dispersion. Dispersion is a concept
commonly discussed relative to optical fiber cable as a limiting factor of
transmission speed. The idea is that travelling through any medium will cause a
small amount of pulse spreading. This means that, if dispersion is strong
enough, the pulse can spread or even break apart into several smaller pulses. To
overcome this, we need enough nonlinearity in the line primarily in terms of
voltage/amplitude dependence such that the wave remains narrow.

@@

While nonlinear capacitors have been used in many studies
\cite{FastHighVoltageNLTL} when implementing this design in high power systems,
varactor diodes have the advantage of a larger capacitance range, and allow us
to perform measurements using lower voltages. In addition, the overall design is
simplified with the use of Varactors because only one bias is needed.

In any case, non linear capacitors are often modeled using equation
(\ref{eq:nonlincap}) to describe the voltage dependence of the
capacitance\cite{nikoo2018theory}.

@@

where $C_{o}$ and $V_{o}$ are zero bias capacitance and reference voltage,
respectively.

@@

This equation is used in section \ref{sub:simsRise} indirectly to estimate the
amount of pulse sharpening we can expect for a given step inputl

@@

In order to properly simulate and predict data using these simulations, we
need first to determine how to model the non-ideal parasitics of the line
and particularly the behavior of the chosen varactor diode loading the NLTL.
There are several ways to accomplish this. The most standard and the method
chosen by our group is SPICE modeling. The disadvantage of this route is
that nonlinear behavior can be particularly hard to model with SPICE alone
at higher frequencies and voltage levels (find source for this, think it's
in SKYWORKS application note).

@@

When constructing a test fixture for measurements of NLTL's of higher frequency,
we must also take into account the properties of the board we are working with.
For our low frequency model and with limited access to robust VNA's this
procedure was not neccessary but we recommend using the two-line pencil
method to approximate the board permittivity and loss tangent as utilized to do
the same for flex PCB's due to it's simplicity and limited need for
equipment\cite{TwoLine}.

For our board, we used a simpler method which may prove useful as a quick and
dirty alternative to the one descbribed above. To find the permitivittivy of the
board we used a simple LCR Meter (DE-5000) and the following formula of a
parallel plate capacitor as shown in equation (\ref{eq:epsAOverD}).

@@

Where we treat the width of the dielectric as plate spacing d, the board area
as area, A, and measured capacitance as C. Solving for permittivity
$\epsilon$, we found a relative permittivity of $\epsilon_{r}\approx 3.54$.
For this measurement a figure normally given for FR4 is roughly 4.0 to 4.4. A
few possible things could cause this error.

The device used to measure the FR4 board generates a test frequency of 100kHz at
maximum. Relative permittivity is usually given for FR4 at a much higher
frequencies in the Mhz range. Another is the precision achievable when measuring
the dielectric thickness. A caliper would be much better suited for this task.
Alternative methods do exist and are quite simple but do necessitate access to a
VNA\cite{SimpleRelative}.

@@

SPICE models are typically easier to find and use in most circuit simulation
software and what we used for our low frequency prototype as it seemed to be
sufficient for our purposes. We chose to use software from the company
SpectrumSoft. We chose this tool due to it's vast array of built-in SPICE models
and capabilities. For example, it ships with SPICE models for the BBY40 and
SMV1249 varactor diodes. Previously a software found in industry for a high
price, SpectrumSoft is now no longer in business and has released this program
free. It can be downloaded from SpectrumSoft's home page.

@@

We were unable to utilize this modeling method. In section \ref{sec:NextSteps}, 
we go over some potentially useful references to help you create models for
higher frequency/bandwidth designs. Most verilog modules we have come across
utilize equation \ref{eq:nonlincap} in order to model non-linear capacitors and
varactors.

@@

In this section we will discuss the basics behind catch-up
theory\cite{wilson1991pulse} and how we confirmed this with simulation.
$\Delta$T represents the time delay at the $N^{th}$ node and is found by the
difference period the line cutoff frequency at zero-bias varactor capacitance
and the period of the line cutoff where varactor capacitance if defined by the
step input maximum. This result is multiplied for every ladder section the pulse
passes. Catch-up theory suggests that this delay is accounted for the variation
in phase velocity as the rising pulse effects the varactor capacitance.

@@

In our simulations we use an input step of 5 volts which takes $1\mu s$ to reach
it's final value. Using equation (\ref{DelayApproxCalc}), we calculate a pulse
sharpening of $8.44ns$ at our $10^{th}$ node and $84.37ns$ at our $100^{th}$
node. To make these calculations, we used the CV plots obtained by measuring the
BBY40 with the test fixture discussed in section \ref{subsec:CVCurves}.

@@

For a lossless line, the sharpest pulse we can achieve can be estimated by first
using equation \ref{NLTLCutoff}\cite{wilson1991pulse}.
We notice that this utilizes the equation of our estimated bragg frequency as
seen in (\ref{eq:bragg}) and describes the differenece in cutoff
frequencies at maximim and minimum bias voltage multiplied by the number of
sections, N.

@@

In figure \ref{fig:riseWithLoss}, we add the series DC resistance of the
inductor $(R_s \approx 122m\Omega)$ and the $1m\Omega$ resistance of the BBY40. 
From this, we see that not much has changed in this model and in fact we see an
increased sharpening effect if anything. We do notice however, that the
amplitude of the wave is much more attenuated at our $100^{th}$ node. This would
lead us to expect a decrease in the pulse sharpening in the circuit instead of a
relatively constant value due to the lessened effect on varactor capacitance as
the pulse climbs.

@@

In order to test all these simulated values we designed a discrete version of
our transmission line using Altium Designer. We emphasized traces being as short
as possible and used Altium's built-in impedance profile tool to match our path
between each pair of SMA connectors to $50\Omega$. The circuit we constructed is
shown assembled in figure \ref{fig:assembled}.

@@

To measure and find the capacitance vs Voltage of our chosen varactors we
created a small ugly style circuit based on a design by W2AEW. Two 100k
resistors are added to limit current flowing to the power supply. A ceramic
capacitor of value much greater than our varactor is placed to protect the meter
from excessive voltage. Since this larger capacitance is in effectively in
parallel with our varactor, our meter measurement is approximately only the
capacitance of the varactor diode itself.

@@

Through use of this fixture, we were able to generate the curves shown in
figures \ref{fig:BBY40CV} and \ref{fig:SMV1249CV}. These were in close agreement
to curves found on datasheets for the BBY40 and SMV1249 respectively.

@@

The frequency response of our circuit effectively tells us the operating
bandwidth of our ciruit. With a 10 ladder long circuit, our cuttoff can be
calculated to be around 103Mhz. We also notice that as we near the corner, the
nonlinearities present in the circuit cause increased amounts of resonance as
expected. The thing to notice is the presence of multiple oscillations of
increasing frequency. These are frequencies we may want to consider when
measuring circuit quality factor, Q.

@@

We had great difficulty in measuring rise time of our circuit and the main
problems we faced were not understanding the theory well enough have do proper
testing. While our components were chosen on the basis of working at 144Mhz, we
failed to see that driving our circuit below this bandwidth and performing
similar simulations may be adequate to prove rise time performance increase.
Another complication was keeping in mind the biasing conditions of the
varactors. We did not design for a port in which to introduce a dc bias to our
circuit and relied on equipment being able to accomplish this. Care must be
taken when using analog generators to supply the DC bias to the circuit. Our
particular model begins sweeping with a negative bias before reaching a positive
one. This effectively puts the varactors into forward bias and allows them to
hold a charge.

Figure \ref{fig:SetupWBiasing} shows the lab setup used to take the time-domain
measurements. A bias-Tee is used to provide a DC bias to our NLTL and is
adjusted by a bench power supply. A picosend pulse generator is used to send
in a well-behaved high speed signal into the DUT.

The input wave is a 1Mhz 1Vpp Square wave with 10nS rise and fall times and the
bias is adjusted from 0 to 3V to further investigate its effect on the
capacitance of the varactor.

@@

Figure \ref{fig:probedNoResistance} shows the expected rise time improvement for
a circuit like ours with 240 ladders sections in an ideal environment. Once we
account for the series resistance, $Rser = -0.7\Omega$ for the purposes of
figure \ref{fig:probedWithResistance}. We see in this case the loss is great
enough to overcome the linearity of the line and we see a rise time increase
instead.

@@

We now change the stimulus to a 1Vpp Sine wave at 1Mhz, and find the power
spectrum the NLTL output. In figure \ref{fig:SimSpec}, it is shows that as a non-linear
circuit there is a great deal of harmonic content generated due to
non-linearity.We expect to see odd and even harmonics throughout the entire
sweep.

Figure \ref{fig:MeasSpec} shows that in the lab, we don't see this same
behavior. This is puzzling and implies that either linear parasitic capacitance
is the dominant factor as a result of either the connectors, board material or
inductor contstruction, or there is enough loss in the line to linearize the
structure or both.

@@

In figure \ref{fig:sparamSchem} shows the schematic used to measure and compare
our measured s-parameters taken through a NanoVNA to our simulated circuit with
parasitics included. The NanoVNA data is imported through use of the n-port
components and define statements are utilized to calculate other parameters such
as generated power in dbm, VSWR, etc...

@@

To produce figure \ref{fig:S11_S21}, we parameterized the values of our
parasitic elements in circuit and within the BBY40 model itself. We changed
these values to match our measured data more closely to see the accuracy of our
model. While we were able to achieve relatively agreeing data we still see a
large discrepancy in our S11 data. At 10MHz, we see a our measured data is
roughly $-10dbm$ lower than our expected simulation. This could be due to
incorrect modeling of our load impedances and requires further study. Some have
mentioned that when using NLTL's, there is bound to always be some impedance
mismatch due to the changing impedance caused by varactor capacitance
\cite{wilson1991pulse}. This may be the case here, but the issue is worth
further simulation.

@@

In figure \ref{fig:VSWRandReturnLoss}, we see that our measured and simulated
are almost identical except for a sharp rise in VSWR near the resonant frequency
of our LC network (144Mhz). After this point, we expect our VSWR to start rising
dramatically. What we see is a much more subtle increase from about 1.5 to 5.
Still a much larger ratio than was present within our operating bandwidth, but
certainly not the expected value of $\approx 100$.

@@

To estimate our circuit Q, we employ a technique called ring-down method
\cite{giangrandi}. Driving the circuit with a square wave much slower than the
expected resonance frequency of the LC network, we simply count the amount of
cycles it takes for the signal envelope to decay to rougly $50\%$ of the maximum
amplitude. Observing figure \ref{fig:EstimQ}, we see that the ringing dies
down in about 1 cycle. To get our Q estimate, we simply multiply this value by
approximate 4.53. This yields an approximate Q of 4.53 for our transmission
line. This seems a little lower than expected however, we can see that the
ringing is not as clean as we expected.
It is unsurprising that our non-linear circuit results in a fluctuating
impedance on the line causing reflections and is prone to generating harmonic
content.

By feeding in a 1Vpp Sine wave and measuring the frequency response of our
circuit, we see that many harmonics should be present in our line. Since we are
using this simulation to stand in place of a physical measurement, this
interference makes sense.

@@

figure \ref{fig:WeirdGeo} shows a possible implementation of pulse sharpening by
taking advantage of the circuit's physical geometry to produce further
non-linearities and mismatches in the circuit. This type of geometry should be
used with much higher frequencies than the designs we created for this report as
it relies on the the lines themselves to act as inductive components. This 
exploration may have several advantages including easier reproduction and less
resistive losses on the line due to lack of discrete inductors.

@@

In this paper, we have summarized the work we have done in understanding the use
of NLTL's and nonlinear waves in general and attempted to reproduce these
results using simulation and measurement. While it was a good first attempt,
there are many areas to explore further such as the simulation, building and
measurement of a model dealing with higher bandwidths (0-6GHz), Non-uniform
geometry, simultaneous edge sharpening, and writing a more accurate Verilog-A
module. For much higher frequencies,the design may need to be constructing using
ADS and momentum as this neccesitates much smaller components and boards.

>>>>

First, we discuss the necessity of non-linear components in such a Nonlinear
Transmission line or \emph{NLTL} for short. We then discuss ways in which the
devices we use in our simulations are modeled, and the obstacles we faced in
constructing these models. We simulate these nonlinear lines using MICROCAP and
discuss possible reasons for the discrepancies we have seen between research and
simulations.

@@

Much like a wave traveling in the ocean towards the shore, certain sections of
the wave travel at different speeds despite the travel of the wave as a whole.
This can be seen first hand by a surfer utilizing their position on the wave to
have the right speed and stability.

In electronics, this same concept is present in optical fibers or waveguides. As
we reflect a signal beam into a guide, the phase velocity can actually exceed
the speed of light, however the speed of the information characterized by
the group velocity is what we care about.

\subsection{Types of NLTL waves}\label{sub:NLTLTypes}
While treating the physics of non-linear waves is beyond the scope of this
paper, the use of NLTL's in transmission lines is founded upon the existence of
waves that do not behave like the pure sine waves we usually consider in
electronics. Both types of waves can be present and used in technology to
lessen the time it takes for an input waveform to reach its final value as well
as lessen the time to return to its ``off'' voltage state. In other words, these
can be used to sharpen both rising and falling edges of a square
wave input\cite{afshari2005nonlinear}.

Informally, \emph{Solitons} are much like the type of wave we see
traveling in the open ocean without breaking and are characterized by their
ability to interact with other waves without losing energy. This means that
after passing another wave it largely maintains its shape and is only shifted in
phase (experiences time delay).

@@

Another main concept of operation of these types of circuits is the need for
non-linearity to outweigh possible wave dispersion. Dispersion is a concept
commonly discussed relative to optical fiber cable as a limiting factor of
transmission speed. The idea is that traveling through any medium will cause a
small amount of pulse spreading. This means that, if dispersion is strong
enough, the pulse can spread or even break apart into several smaller pulses. To
overcome this, we need enough nonlinearity in the line primarily in terms of
voltage/amplitude dependence such that the wave remains narrow.

@@

While nonlinear capacitors have been used in many studies
\cite{FastHighVoltageNLTL} when implementing this design in high power systems,
varactor diodes have the advantage of a larger capacitance range and allow us
to perform measurements using lower voltages. In addition, the overall design is
simplified with the use of Varactors because only one bias is needed.

In any case, nonlinear capacitors are often modeled using equation
(\ref{eq:nonlincap}) to describe the voltage dependence of the
capacitance\cite{nikoo2018theory}.

@@

where $C_{o}$ and $V_{o}$ are zero bias capacitance and the reference voltage,
respectively.

@@

This equation is used in section \ref{sub:simsRise} indirectly to estimate the
amount of pulse sharpening we can expect for a given step input

@@

In order to properly simulate and predict data using these simulations, we
need first to determine how to model the non-ideal parasitics of the line
and particularly the behavior of the chosen varactor diode loading the NLTL.
There are several ways to accomplish this. The most standard and the method
chosen by our group is SPICE modeling. The disadvantage of this route is
that nonlinear behavior can be particularly hard to model with SPICE alone
at higher frequencies and voltage levels (find a source for this, think it's
in SKYWORKS application note).

@@

When constructing a test fixture for measurements of NLTL's of higher frequency,
we must also take into account the properties of the board we are working with.
For our low-frequency model and with limited access to robust VNA's, the
the procedure was not necessary but we recommend using the two-line pencil
method to approximate the board permittivity and loss tangent as utilized to do
the same for flex PCB's due to its simplicity and limited need for
equipment\cite{TwoLine}.

For our board, we used a simpler method which may prove useful as a quick and
dirty alternative to the one described above. To find the permittivity of the
board we used a simple LCR Meter (DE-5000) and the following formula of a
parallel plate capacitor as shown in equation (\ref{eq:epsAOverD}).

@@

Where we treat the width of the dielectric as plate spacing d, the board area
as area, A, and measured capacitance as C. Solving for permittivity
$\epsilon$, we found a relative permittivity of $\epsilon_{r}\approx 3.54$.
For this measurement, a figure normally given for FR4 is roughly 4.0 to 4.4. A
few possible things could cause this error.

The device used to measure the FR4 board generates a test frequency of 100kHz at
maximum. Relative permittivity is usually given for FR4 at much higher
frequencies in the Mhz range. Another is the precision achievable when measuring
the dielectric thickness. A caliper would be much better suited for this task.
Alternative methods do exist and are quite simple but do necessitate access to a
VNA\cite{SimpleRelative}.

@@

SPICE models are typically easier to find and use in most circuit simulation
software and what we used for our low-frequency prototype as it seemed to be
sufficient for our purposes. We chose to use software from the company
SpectrumSoft. We chose this tool due to its vast array of built-in SPICE models
and capabilities. For example, it ships with SPICE models for the BBY40 and
SMV1249 varactor diodes. Previously a software found in the industry for a high
price, SpectrumSoft is now no longer in business and has released this program
free. It can be downloaded from SpectrumSoft's home page.

@@

We were unable to utilize this modeling method. In section \ref{sec:NextSteps}, 
we go over some potentially useful references to help you create models for
higher frequency/bandwidth designs. Most Verilog modules we have come across
utilize equation \ref{eq:nonlincap} in order to model non-linear capacitors and
varactors.

@@

In this section, we will discuss the basics behind catch-up
theory\cite{wilson1991pulse} and how we confirmed this with simulation.
$\Delta$T represents the time delay at the $N^{th}$ node and is found by the
difference period the line cutoff frequency at zero-bias varactor capacitance
and the period of the line cutoff where varactor capacitance is defined by the
step input maximum. This result is multiplied for every ladder section the pulse
passes. Catch-up theory suggests that this delay is accounted for the variation
in phase velocity as the rising pulse effects the varactor capacitance.

@@

In our simulations, we use an input step of 5 volts which takes $1\mu s$ to reach
its final value. Using equation (\ref{DelayApproxCalc}), we calculate a pulse
sharpening of $8.44ns$ at our $10^{th}$ node and $84.37ns$ at our $100^{th}$
node. To make these calculations, we used the CV plots obtained by measuring the
BBY40 with the test fixture discussed in section \ref{subsec:CVCurves}.

@@

For a lossless line, the sharpest pulse we can achieve can be estimated by first
using equation \ref{NLTLCutoff}\cite{wilson1991pulse}.
We notice that this utilizes the equation of our estimated Bragg frequency as
seen in (\ref{eq:bragg}) and describes the difference in cutoff
frequencies at maximum and minimum bias voltage multiplied by the number of
sections, N.

@@

In figure \ref{fig:riseWithLoss}, we add the series DC resistance of the
inductor $(R_s \approx 122m\Omega)$ and the $1m\Omega$ resistance of the BBY40. 
From this, we see that not much has changed in this model and in fact, we see an
increased sharpening effect if anything. We do notice, however, that the
amplitude of the wave is much more attenuated at our $100^{th}$ node. This would
lead us to expect a decrease in the pulse sharpening in the circuit instead of a
relatively constant value due to the lessened effect on varactor capacitance as
the pulse climbs.

@@

In order to test all these simulated values, we designed a discrete version of
our transmission line using Altium Designer. We emphasized traces being as short
as possible and used Altium's built-in impedance profile tool to match our path
between each pair of SMA connectors to $50\Omega$. The circuit we constructed is
shown assembled in figure \ref{fig:assembled}.

@@

To measure and find the capacitance vs Voltage of our chosen varactors we
created a small ugly style circuit based on a design by W2AEW. Two 100k
resistors are added to limit the current flowing to the power supply. A ceramic
capacitor of value much greater than our varactor is placed to protect the meter
from excessive voltage. Since this larger capacitance effectively in
parallel with our varactor, our meter measurement is approximately only the
capacitance of the varactor diode itself.

@@

Through the use of this fixture, we were able to generate the curves shown in
figures \ref{fig:BBY40CV} and \ref{fig:SMV1249CV}. These were in close agreement
to curves found on datasheets for the BBY40 and SMV1249 respectively.

@@

The frequency response of our circuit effectively tells us the operating
bandwidth of our circuit. With a 10 ladder long circuit, our cutoff can be
calculated to be around 103Mhz. We also notice that as we near the corner, the
nonlinearities present in the circuit cause increased amounts of resonance as
expected. The thing to notice is the presence of multiple oscillations of
increasing frequency. These are frequencies we may want to consider when
measuring circuit quality factor, Q.

@@

We had great difficulty in measuring the rise time of our circuit and the main
problems we faced were not understanding the theory well enough have done proper
testing. While our components were chosen on the basis of working at 144Mhz, we
failed to see that driving our circuit below this bandwidth and performing
similar simulations may be adequate to prove rise time performance increase.
Another complication was keeping in mind the biasing conditions of the
varactors. We did not design for a port in which to introduce a dc bias to our
circuit and relied on equipment being able to accomplish this. Care must be
taken when using analog generators to supply the DC bias to the circuit. Our
particular model begins sweeping with a negative bias before reaching a positive
one. This effectively puts the varactors into forward bias and allows them to
hold a charge.

Figure \ref{fig:SetupWBiasing} shows the lab setup used to take the time-domain
measurements. A bias-Tee is used to provide a DC bias to our NLTL and is
adjusted by a bench power supply. A picosecond pulse generator is used to send
in a well-behaved high-speed signal into the DUT.

The input wave is a 1Mhz 1Vpp Square wave with 10nS rise and fall times and the
bias is adjusted from 0 to 3V to further investigate its effect on the
the capacitance of the varactor.

@@

Figure \ref{fig:probedNoResistance} shows the expected rise time improvement for
a circuit like ours with 240 ladders sections in an ideal environment. Once we
account for the series resistance, $R_{ser} = 0.7\Omega$ for the purposes of
figure \ref{fig:probedWithResistance}. We see in this case the loss is great
enough to overcome the linearity of the line and we see a rise time increase
instead.

@@

We now change the stimulus to a 1Vpp Sine wave at 1Mhz and find the power
spectrum of the NLTL output. In figure \ref{fig:SimSpec}, it is shows that as a non-linear
circuit there is a great deal of harmonic content generated due to
non-linearity. We expect to see odd and even harmonics throughout the entire
sweep.

Figure \ref{fig:MeasSpec} shows that in the lab, we don't see this same
behavior. This is puzzling and implies that either linear parasitic capacitance
is the dominant factor as a result of either the connectors, board material or
inductor construction or there is enough loss in the line to linearize the
structure or both.

@@

In figure \ref{fig:sparamSchem} shows the schematic used to measure and compare
our measured s-parameters. These were measured with a NanoVNA. The circuit simulates included parasitics. The NanoVNA data is imported through the use of n-port
components and define statements are utilized to calculate other parameters such
as generated power in dBm, VSWR, etc...

@@

To produce figure \ref{fig:S11_S21}, we parameterized the values of our
parasitic elements in the circuit and within the BBY40 model itself. We changed
these values to match our measured data more closely to see the accuracy of our
model. While we were able to achieve relatively agreeing data we still see a
large discrepancy in our S11 data. At 10MHz, we see our measured data is
roughly $-10dbm$ lower than our expected simulation. This could be due to
incorrect modeling of our load impedances and requires further study. Some have
mentioned that when using NLTL's, there is bound to always be some impedance
mismatch due to the changing impedance caused by varactor capacitance
\cite{wilson1991pulse}. This may be the case here, but the issue is worth
further simulation.

@@

In figure \ref{fig:VSWRandReturnLoss}, we see that our measured and simulated
are almost identical except for a sharp rise in VSWR near the resonant frequency
of our LC network (144Mhz). After this point, we expect our VSWR to start rising
dramatically. What we see is a much more subtle increase from about 1.5 to 5.
Still, a much larger ratio than was present within our operating bandwidth, but
certainly not the expected value of $\approx 100$.

@@

To estimate our circuit Q, we employ a technique called ring-down method
\cite{giangrandi}. Driving the circuit with a square wave much slower than the
expected resonance frequency of the LC network, we simply count the number of
cycles it takes for the signal envelope to decay to roughly $50\%$ of the maximum
amplitude. Observing figure \ref{fig:EstimQ}, we see that the ringing dies
down in about 1 cycle. To get our Q estimate, we simply multiply this value by 4.53. This yields an approximate Q of 4.53 for our transmission
line. This seems a little lower than expected however, we can see that the
ringing is not as clean as we expected.
It is unsurprising that our non-linear circuit results in a fluctuating
impedance on the line causing reflections and is prone to generating harmonic
content.

By feeding in a 1Vpp Sine wave and measuring the frequency response of our
circuit, we see that many harmonics should be present in our line. Since we are
using this simulation to stand in place of physical measurement, this
interference makes sense.

@@

figure \ref{fig:WeirdGeo} shows a possible implementation of pulse sharpening by
taking advantage of the circuit's physical geometry to produce further
non-linearities and mismatches in the circuit. This type of geometry should be
used with much higher frequencies than the designs we created for this report as
it relies on the lines themselves to act as inductive components. This exploration may have several advantages including easier reproduction and less
resistive losses on the line due to lack of discrete inductors.

@@

In this paper, we have summarized the work we have done in understanding the use
of NLTL's and nonlinear waves in general and attempted to reproduce these
results using simulation and measurement. While it was a good first attempt,
there are many areas to explore further such as the simulation, building and
measurement of a model dealing with higher bandwidths (0-6GHz), Non-uniform
geometry, simultaneous edge sharpening, and writing a more accurate Verilog-A
module. For much higher frequencies, the design may need to be constructing using
ADS and momentum as this necessitates much smaller components and boards.
